---
title: "babynames post"
description: |
  play with baby names data 
author: "Jenny Richmond"
date: 01-6-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# data viz with ggplot


### load packages 

We can use the `library()` function to load the packages we need. The `tidyverse` package contains tons of useful functions for data wrangling and visualisation (including ggplot). The `ozbabynames` package contains data from birth records in Australia. 


```{r}
library(tidyverse)
library(ozbabynames)

```

### read the babynames data

```{r}
ozbabynames <- ozbabynames

glimpse(ozbabynames)

```


### define the names you want to compare

btw Lady Gaga's real name is Stefani

```{r}
artist_names <- c("Billie", "Taylor", "Stefani")

```

### whose name is most popular?? 

```{r}

ozbabynames %>%
  filter(name %in% artist_names) %>%
  group_by(name, year) %>% 
  summarise(count = sum(count)) %>%
  ggplot(aes(x = year, 
             y = count,
             colour = name)) +
  geom_line() +
  theme_bw() +
 facet_wrap(~name,
             scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "artist names plot with free_y")
```



What does the `scales = "free y"` do? What happens if you delete it? Is the plot more meaningful with "free y" or without it?

```{r}
ozbabynames %>%
  filter(name %in% artist_names) %>%
  group_by(name, year) %>% 
  summarise(count = sum(count)) %>%
  ggplot(aes(x = year, 
             y = count,
             colour = name)) +
  geom_line() +
  theme_bw() +
 facet_wrap(~name) +
  theme(legend.position = "none") +
  labs(title = "artist names plot without free_y")
```



### Recreate the plot above with your names 

```{r}
your_names <- c("Jenny", "Kate", "Danielle")
```

What do you need to change about the code below to make it plot your names?

```{r}
ozbabynames %>%
  filter(name %in% your_names) %>%
  group_by(name, year) %>% 
  summarise(count = sum(count)) %>%
  ggplot(aes(x = year, 
             y = count,
             colour = name)) +
  geom_line() +
  theme_bw() +
 facet_wrap(~name,
             scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "our names plot with free_y")
```

Do you need `scales = "free_y"`?

```{r}
ozbabynames %>%
  filter(name %in% your_names) %>%
  group_by(name, year) %>% 
  summarise(count = sum(count)) %>%
  ggplot(aes(x = year, 
             y = count,
             colour = name)) +
  geom_line() +
  theme_bw() +
 facet_wrap(~name) +
  theme(legend.position = "none") +
    labs(title = "our names plot without free_y")
```



